% Updated abstract: interim MAEG4998/4999 PDF (Jan 2026) + repository updates (2026)
\subsection*{Main objectives}
This project aims to develop a software-centric enhancement module for a lower-limb robotic exoskeleton designed to assist workers in heavy load transportation tasks ($>20$~kg) in industrial, logistics, and construction environments. The primary objectives are to enable real-time detection of user locomotion intent (including walking, stair ascent/descent, sit-to-stand transitions, and speed variations), accurately estimate joint kinematics through multi-sensor fusion, and deliver adaptive personalized torque assistance to minimize physical effort and injury risk. To achieve this, the module will incorporate computational techniques to handle variability in user biomechanics and environmental conditions, ensuring robust performance across diverse scenarios.

\subsection*{Planned methods (interim report, Jan 2026 PDF)}
The proposed system integrates inertial measurement units (IMUs) placed on the lower limbs and back, with data processed through a Kalman filter-based sensor fusion pipeline implemented in MATLAB to achieve robust, low-latency joint angle and velocity estimation. Locomotion mode and transition prediction were planned using machine learning models (Support Vector Machines and Long Short-Term Memory networks) trained on labeled gait datasets, targeting $>90$\% classification accuracy and a 100--200~ms predictive horizon, with performance evaluated via confusion matrices, precision, and recall. Finite state machines were incorporated for locomotion mode transitions. The data collection protocol involves attaching three mobile phones as IMUs to the hips and lower back; performing standardized tasks such as walking 10~m on flat ground, climbing 10 steps, or navigating uneven surfaces at self-selected speeds; and exporting accelerometer, gyroscope, and magnetometer data as CSV files for processing. Joint kinematics information is intended for precise control.

\subsection*{Implemented methods (repository baseline, updated)}
The validated MATLAB baseline uses overlapping windows (100 samples at 100~Hz, step 50). Kalman fusion (\texttt{imufilter}) is retained for simulation-side kinematic visualization, not as mandatory preprocessing for the classifier.

The \textbf{primary deployment classifier} is a \textbf{binary} RBF SVM (\texttt{fitcsvm}, standardized features) with a \textbf{30-dimensional} training vector: five statistics per IMU slot $\times$ six slots (\texttt{Features.m}). The final project pipeline is centered on \textbf{HuGaDB}, while pseudo-real-time replay is kept as a separate simulation workflow. USC-HAD remains in the repository for benchmark comparison under the same feature layout. The positive class groups \textbf{active-movement} activities per dataset-specific label maps.

A \textbf{multiclass} ECOC ensemble (\texttt{fitcecoc}) provides \textbf{dataset-native} activity labels for analysis (twelve USC-HAD trial classes; twelve HuGaDB sample classes); exoskeleton gating uses the binary model and an asymmetric FSM (3 consecutive active predictions to assist, 5 to release).

An \textbf{optional binary LSTM} (Deep Learning Toolbox) classifies the same active-movement vs.\ inactive task from \textbf{sequence} inputs (raw IMU windows mapped to feature$\times$time matrices) via \texttt{TrainLstmBinary} and \texttt{RunExoskeletonPipelineLstm}; it is an alternative to the RBF SVM, not a hybrid SVM+LSTM fusion stage.

The 100--200~ms horizon remains a \textit{design target}; literature reports comparable latencies for kinematic lift-onset cues~\cite{chen2018}, but the submitted baselines were not benchmarked under that specific end-to-end metric.

\subsection*{Interim results statement (Dec 2025 / Jan 2026 PDF)}
The interim submission reported: a functional three-phone acquisition prototype; preliminary Kalman filtering and intention-detection algorithms validated on public datasets including HuGaDB and the USC-SIPI gait database; repository setup; real-time visualization; and initial locomotion classification with \textbf{98.88\%} accuracy and a \textbf{150~ms} predictive horizon. \textit{Corrections:} the smartphone gait database used in this project is \textbf{USC-HAD}~\cite{zhang2012}, not USC-SIPI. The 150~ms figure overlapped claims from the literature on hip-based lift detection~\cite{chen2018} and was not rigorously established as an end-to-end latency of the final classifier at submission time.

\subsection*{Updated empirical results (per-dataset public-data metrics)}
Kalman-filtered visualization and SVM-based intention detection are implemented in MATLAB. \textbf{Binary SVM} five-fold out-of-fold accuracy from the benchmark evaluation files is \textbf{\BinaryOofAccUsc\%} (USC-HAD) and \textbf{\BinaryOofAccHuGa\%} (HuGaDB). Native ECOC multiclass SVM accuracies on stratified subsamples are \textbf{\MulticlassOofAccUsc\%} (USC-HAD) and \textbf{\MulticlassOofAccHuGa\%} (HuGaDB).

\ifnum\HasBinaryLstmMetrics=1
Binary LSTM holdout accuracies are \textbf{\LstmBinaryHoldoutAccUsc\%} (USC-HAD) and \textbf{\LstmBinaryHoldoutAccHuGa\%} (HuGaDB).
\else
Binary LSTM metrics are optional and are exported after running the dataset-specific LSTM evaluation scripts.
\fi
\ifnum\HasMulticlassLstmMetrics=1
 Native multiclass LSTM holdout accuracies are \textbf{\LstmMulticlassHoldoutAccUsc\%} (USC-HAD) and \textbf{\LstmMulticlassHoldoutAccHuGa\%} (HuGaDB).
\else
 Multiclass LSTM metrics are optional and are exported after running the new per-dataset sequence models.
\fi
Real-time pipeline simulation replays held-out HuGaDB sessions directly and shows Kalman-smoothed segment pitch together with FSM-latched assist commands. Scalar accuracies in this report are regenerated into \texttt{docs/latex/generated\_metrics.tex} by \texttt{scripts/ExportMetricsForReport.m}.

\subsection*{Main conclusions}
This project demonstrates that automation technology, particularly machine-learning-based intention recognition combined with sensor fusion and finite state machines, is applicable to next-generation powered exoskeleton control supervisors. When fully realized, the proposed module has strong potential for integration into industrial platforms (e.g., ExoTechHK Limited), offering a reproducible HuGaDB-based assist-gating baseline while preserving per-dataset comparison evidence in the write-up. This work contributes to workplace ergonomics and provides a reproducible foundation for hardware-in-the-loop studies. The repository now documents both the \textbf{original project plan} (hybrid SVM+LSTM fusion, extended fusion) and the \textbf{delivered MATLAB baseline} (HuGaDB deployment path, supporting per-dataset SVM/LSTM comparisons, optional multiclass SVM, and per-dataset multiclass LSTM sequence baselines).
