% Chapter 5: summary + interim PDF remaining work + timeline + risks + repository status
\section{Summary of what was delivered}
This project produced a reproducible MATLAB baseline for lower-limb exoskeleton \emph{assist gating} on public IMU data: Kalman fusion (\texttt{imufilter}), a 30-dimensional feature pipeline with merged USC-HAD and HuGaDB training, binary RBF SVM classification with five-fold out-of-fold evaluation, an optional twelve-class ECOC model for exploratory activity recognition, an optional binary LSTM sequence classifier, and finite-state-machine smoothing with asymmetric hysteresis for stable on/off assist commands in simulation. Scalar accuracies and figures are traceable to versioned scripts and \texttt{results/*.mat}, with numbers exported into \texttt{generated\_metrics.tex} via \texttt{ExportMetricsForReport.m}.

The binary accuracy target ($>90$\%) is met for USC-HAD-only and merged training; HuGaDB-only accuracy is lower, consistent with dataset heterogeneity and label mapping choices. Multiclass accuracy remains modest and is not proposed as a deployable control layer without further work. End-to-end hardware trials, embedded deployment, hybrid SVM--LSTM fusion as in the original pseudocode, and the full predictive-horizon benchmark were not completed in software; the sections below preserve the interim roadmap, risks, and timeline so planned scope stays distinguishable from the validated repository.

\section{Remaining tasks and roadmap (updated for final report)}
The Jan~2026 interim plan listed scope still open beyond the validated MATLAB baseline. Relative to the repository status in Section~\ref{sec:repo_status} and the results chapter, the backlog can be restated as follows:
\begin{itemize}
  \item \textbf{Intention detection:} Binary SVM and optional LSTM/multiclass paths exist in-repo; remaining work includes harmonizing LSTM evaluation with the five-fold OOF protocol used for SVMs, optional SVM--LSTM or ensemble fusion (not implemented), richer transition handling (multi-mode, speed), and deployment-time tuning on partner or collected data.
  \item \textbf{Machine learning deployment:} On-device inference from trained SVM/LSTM artifacts (e.g., MATLAB Coder or an embedded runtime), plus any terrain- or activity-specific models beyond the current public-dataset baselines.
  \item \textbf{Hardware integration:} Port validated decision logic to the ExoTechHK controller stack and close the loop with actuation, beyond MATLAB simulation and CSV replay.
  \item \textbf{Testing and validation:} Instrumented bench tests, ethical user trials under torque limits, and field-style evaluation (latency, confusion-matrix metrics, precision/recall under realistic loads).
  \item \textbf{Documentation and presentation:} This document consolidates the technical baseline; slides, video demonstration, and comparative benchmarks against additional literature or commercial systems remain optional deliverables beyond the core report.
\end{itemize}

\section{Future work beyond the FYP (interim PDF)}
\begin{itemize}
  \item \textbf{Personalization:} Short calibration walks with transfer learning and lightweight RL (e.g., PPO/SAC) to tune assistance.
  \item \textbf{Injury-risk assessment:} Integrate ergonomic risk frameworks such as Zelik et al.~\cite{zelik2022} into user trials.
\end{itemize}

\section{Project timeline (interim PDF)}
Table~\ref{tab:timeline} reproduces the phased plan from the Jan 2026 report (notional dates; confirm with your advisor).

\begin{table}[htbp]
  \centering
  \small
  \caption{Phased development plan (interim submission).}
  \label{tab:timeline}
  \begin{tabular}{@{}p{2.7cm}p{2.3cm}p{4.4cm}p{4.4cm}@{}}
    \toprule
    \textbf{Phase} & \textbf{Duration} & \textbf{Key activities} & \textbf{Milestones} \\
    \midrule
    ML training \& simulation integration & Jan -- Feb 2026 (2 months) & Train/deploy ML; Simulink HIL-style virtual tests; locomotion modes under load & Validated models ($\geq 90$\% target); simulation demo \\
    Hardware integration \& testing & Mar -- Apr 15, 2026 (1.5 months) & Software--hardware integration; bench and limited treadmill tests (5--10 trials); ethics & Logs with latency/error metrics \\
    Optimization, documentation, finalization & Apr 16 -- May 2026 (1.5 months) & Refine from tests; compile report; presentation & Submitted report; presentation; archived repository \\
    \bottomrule
  \end{tabular}
\end{table}

\section{Risk assessment (interim PDF)}
Table~\ref{tab:risks} summarizes risks and mitigations from the Jan 2026 report.

\begin{table}[htbp]
  \centering
  \small
  \caption{Risks and mitigations (interim submission).}
  \label{tab:risks}
  \begin{tabular}{@{}p{3.6cm}ccp{5.5cm}@{}}
    \toprule
    \textbf{Risk} & \textbf{Like.} & \textbf{Imp.} & \textbf{Mitigation} \\
    \midrule
    Hardware integration delays (e.g., actuator vs.\ ExoTechHK frame) & M & H & Modular COTS parts; parallel software; early prototype access \\
    ML underperformance ($<85$\% accuracy) & L & M & Rule/ML hybrid fallback; partner datasets; cross-validation / RL tuning \\
    Trial delays & M & H & Public/simulated data interim; limit volunteers; secure IRB early \\
    Budget overrun & L & L & University funding; HKSTP supplements \\
    \bottomrule
  \end{tabular}
\end{table}

\section{Repository status (consolidated, 2026)}\label{sec:repo_status}
Relative to the interim plan, the following are \textbf{delivered in software}: Kalman fusion baseline; binary RBF SVM with merged USC-HAD+HuGaDB training and dataset ablation scripts; optional twelve-class ECOC SVM; optional binary LSTM sequence classifier (\texttt{TrainLstmBinary}, \texttt{EvaluateLstmConfusion}, \texttt{RunExoskeletonPipelineLstm}); asymmetric FSM; documentation and evaluation figures.

\textbf{Not delivered:} hybrid SVM+LSTM fusion as in the original pseudocode, MATLAB Coder deployment, ROS/PID torque layer, hardware controller integration, and prospective user trials---these remain aligned with the roadmap and risk tables above.

\textbf{Synthesis:} The LaTeX report intentionally preserves the \textbf{full interim narrative} (objectives, proposed LSTM/EKF/ROS extensions, timeline, risks) while adding a \textbf{faithful description of the implemented repository}, so readers can distinguish planned scope from validated baseline.
