% Final Year Project Final Report -- LaTeX replica (MAEG4999 style)
% Compile from this directory: pdflatex main.tex (twice for TOC)
\documentclass[12pt,a4paper]{report}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{setspace}
\setstretch{1.15}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{array}
\usepackage{longtable}
\usepackage{enumitem}
\usepackage{amsmath,amssymb}
% IEEE-style numeric citations (compression, sorting); load before hyperref
\usepackage{cite}
% Allow line breaks in long DOI URLs (references.bib url={https://doi.org/...})
\usepackage[hyphens]{url}
\usepackage[hidelinks]{hyperref}

% Figure path: project results folder (compile from docs/latex/)
\graphicspath{{../../results/}}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[C]{\small MAEG4999 Final Year Project II: Robotic Exoskeleton for Load Transportation}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0.4pt}
\fancypagestyle{plain}{%
  \fancyhf{}
  \fancyfoot[C]{\thepage}
  \renewcommand{\headrulewidth}{0pt}
}

\hypersetup{
  pdftitle={Robotic Exoskeleton for Load Transportation - Final Report},
  pdfauthor={Yim Fu Chong}
}

\begin{document}

% ---------- Roman front matter ----------
\pagenumbering{roman}

\begin{titlepage}
  \centering
  \vspace*{1.5cm}
  {\large WHL3.}\\[1.2cm]
  {\LARGE \textbf{Robotic Exoskeleton for Load Transportation}}\\[1.5cm]
  By: Yim Fu Chong (1155176974)\\[0.8cm]
  Supervisor: Professor Wei-Hsin Liao\\[1.5cm]
  {\large Final Year Project Final Report}\\[0.6cm]
  Submitted in Partial Fulfillment of the\\
  Requirements for\\[0.5cm]
  The Degree of Bachelor of Engineering\\[0.8cm]
  Department of Mechanical and Automation Engineering\\
  The Chinese University of Hong Kong\\[1.5cm]
  {\large \today}
  \vfill
\end{titlepage}

\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}
I would like to express my sincere gratitude to my supervisor Professor Wei-Hsin Liao. His insights into robotic exoskeletons and actuation systems have been instrumental in shaping the direction and progress of my work.

I am also deeply thankful to the team at ExoTechHK Limited for their collaboration and providing access to their lower-limb exoskeleton prototype. This partnership, facilitated under the Hong Kong Science and Technology Parks (HKSTP) incubation programme, has offered a practical pathway for technology transfer and real-world application.

Special thanks go to the Department of Mechanical and Automation Engineering at The Chinese University of Hong Kong for the resources and facilities that enabled this research. Additionally, I appreciate the contributions from fellow researchers and peers who provided feedback during the development phase.

Finally, I extend my appreciation to my family and friends for their encouragement and understanding during this challenging yet rewarding endeavor.

\begin{flushright}
Yim Fu Chong\\
\today
\end{flushright}

\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}
\input{chapters/abstract_merged}

\tableofcontents
\clearpage

% ---------- Main body ----------
\pagenumbering{arabic}
\setcounter{page}{1}

\chapter{Problem Description}

\section{Problem Statement}
Manual load transportation in industrial, construction, and logistics settings frequently involves carrying loads exceeding 20~kg, surpassing human physiological limits and leading to high rates of lower-back pain, musculoskeletal disorders (MSDs), reduced productivity, and billions in annual economic costs worldwide. Existing assistive exoskeletons often fail to provide proactive, real-time adaptation to the user's locomotion intent or environmental changes, resulting in delayed or mismatched assistance that compromises safety and comfort. Moreover, generic assistance profiles do not account for inter-individual differences in gait asymmetry, fatigue thresholds, anthropometrics, or preferred movement patterns. These limitations are especially critical in dynamic workplaces where load magnitude, terrain, and tasks constantly vary. There remains a clear need for a lightweight, cost-effective lower-limb exoskeleton that delivers accurate intent detection, robust sensor fusion, and truly personalized adaptive control---without relying on invasive interfaces such as electromyography (EMG) or requiring extensive user training.

\section{Project Goal}
The primary goal is to develop a software-centric enhancement module for ExoTechHK Limited's existing lower-limb exoskeleton prototype---a local startup incubated at HKSTP and founded by CUHK researchers---to augment human strength and endurance during heavy load-carrying tasks ($>20$~kg) without invasive sensors or lengthy calibration.

\subsection*{Original project goal statement (interim PDF, Jan 2026)}
The module was specified to integrate: (i)~a \textbf{data acquisition and processing pipeline} using three mobile phones as IMUs, Kalman-based fusion for joint kinematics, and \textit{initial} machine learning (e.g., hybrid SVM+LSTM) for locomotion classification with validation on public datasets; and (ii)~\textbf{proactive real-time locomotion intent recognition} using hip and lower-back IMU data via a hybrid SVM+LSTM model with multi-sensor fusion, targeting $>90$\% intent-detection accuracy and a 100--200~ms predictive horizon, with evaluation through confusion matrices.

\subsection*{Current implementation emphasis (repository)}
The module integrates two core components as realized in code:
\begin{itemize}[noitemsep]
  \item \textbf{Data Acquisition and Processing Pipeline:} Three-phone CSV prototype; Kalman fusion (\texttt{imufilter}); \textbf{binary} RBF SVM training on merged USC-HAD and HuGaDB (30-D window features) with USC-only / HuGaDB-only ablations; optional \textbf{twelve-class} ECOC SVM; optional \textbf{binary LSTM} on sequence-shaped IMU windows (\texttt{TrainLstmBinary}, \texttt{RunExoskeletonPipelineLstm}). Embedded deployment remains future work.
  \item \textbf{Proactive Real-Time Locomotion Intent Recognition:} Binary task = \textbf{locomotion-like} vs.\ \textbf{non-locomotion} from dataset label maps. The $>90$\% target is met for USC-HAD-only and merged binary SVM training; HuGaDB-only binary accuracy is lower (imbalance and label noise). The optional LSTM uses a 20\% holdout metric (not five-fold OOF). The 100--200~ms horizon is a design target, not re-benchmarked here for the submitted baselines.
\end{itemize}

Being entirely software-based, the solution enables seamless integration into existing hardware prototypes, rapid deployment, and scalability across future platforms.

\section{Significance of the Project}
This project directly addresses a major occupational health challenge affecting industrial workers, logistics and construction personnel, and military operators exposed to heavy manual handling. In Hong Kong, manual handling injuries are a leading cause of absenteeism and healthcare expenditure. By providing proactive, personalized torque assistance at the hips and knees using cooperative control strategies---in which the exoskeleton follows rather than leads the user---the system aims for high accuracy in intent detection; recent field evidence shows that such user-following approaches improve usability compared to fixed profiles~\cite{le2024}.

Conducted in close collaboration with ExoTechHK Limited under the HKSTP incubation programme, the project offers a pathway for technology transfer and commercialization. Software-centric design enables low-cost retrofitting to existing lower-limb platforms and upgrades across hardware generations, with potential to broaden adoption. The technology aligns with United Nations Sustainable Development Goal 8 (Decent Work and Economic Growth) and strengthens Hong Kong's innovation ecosystem through industry--academia partnership. The standardized mobile-phone IMU protocol supports reproducibility and open collaboration in biomechanics and wearable technology.

\input{chapters/literature_review}

\chapter{Proposed Solution Methods}
\input{chapters/methods_merged}

\chapter{Results and Discussions}
\input{chapters/results_merged}

\chapter{Conclusions}
\input{chapters/conclusions_merged}

\clearpage
\input{chapters/references}

\end{document}
